---
title: "Schedule weekly 1-on-1 meetings"
output: 
   github_document:
    html_preview: false
    df_print: kable
    number_sections: yes
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

This script tries to schedule weekly 1-1 meeting based on a presumed data format (a sign-up sheet in a special format). We will read this data from a Google spreadsheet. Presumably, we want different students to sign up on this spreadsheet. We want to output a list of 1-on-1 meeting slots.

Our example is at https://docs.google.com/spreadsheets/d/1KYSRe7Wjk7eMQ8e5zqr4U2Y1iC-FKI00Hk8UydrRc6g/edit?usp=sharing (what's the md syntax of link?). Each student, for special reasons, is called a duck 🦆. The big duck is the advisor. Technically we can exclude the big duck column; but to be able to reuse this code in the future, we put the advisor in a separate column.

```{r message=FALSE}
library(googlesheets4, quietly = TRUE)
library(magrittr, quietly = TRUE) # %T>% 
library(tidyverse, quietly = TRUE)
library(ascii, quietly = TRUE) # print table in plain text
gs4_deauth()
```

# Read data

First, we read the signup sheet. To make our life easier, we just read the specified range.

```{r}

signup_raw = range_read("https://docs.google.com/spreadsheets/d/1KYSRe7Wjk7eMQ8e5zqr4U2Y1iC-FKI00Hk8UydrRc6g/edit?usp=sharing",
           range = 'A1:M49',
           sheet = 'test'
           ) 

```

Sanity check

```{r}
head(signup_raw)
```

# Format data

Let's first transform data into a format that's easier to work with.

```{r}
signup <- 
signup_raw %>% 
 mutate(across(everything(), ~replace(., . ==  'Yes' , 2))) %>% 
 mutate(across(everything(), ~replace(., . ==  'No' , 0))) %>% 
 mutate(across(everything(), ~replace(., . ==  'Maybe' , 1))) %>% 
 # transform each column into a number 
 # this assumes the 3 to last column is signups 
 mutate(across(3:length(.), as.numeric)) %>% 
 mutate(
   # some formatting tricks
   daytrick = as.numeric(as.factor(day)) * 100, 
   timeslot = daytrick + as.numeric(as.factor(time))) %>% 
 relocate(timeslot, .after = 'time') %>% 
  # rm columns which we don't need
 select(-daytrick)


```


```{r}
head(signup)
```


# Compute availabilities

we compute each duck's availability shared with the big duck.

```{r}
duck_avails <- 
signup %>% 
  select(-time) %>% 
  # if the duck shares the slot with the big duck
  # (maybe) we can constrain from the big duck?
  mutate(across(4:length(.), ~ (. > 0 & bigduck > 0))) %>% 
  #print() %>%
  # and if the duck is available after this half an hour slot
  mutate(across(4:length(.), ~ (. & lead(.)))) %>% 
  #print() %>%
  mutate_at(4:length(.), ~replace_na(.,FALSE))

head(duck_avails)
```


# Compute heuristics

We then compute heuristics based on the number of possible arrangements

```{r}
duck_avail_heuristics <- duck_avails %>% 
  select(4:length(.)) %>% 
  # separate days
  split(., as.factor(signup$day)) %>% 
  # the possible arranges on each day
  # remove the first one because 
  lapply(\(x) sapply(x, sum))  %T>% 
  # print each duck's possible arranges on each day
  print() %>%
  # bind them together
  abind::abind(along = 0) %>% 
  as_tibble() %>% 
  # sort
  sapply(\(x) sum(x)) %>% sort()
```

how many possible arranges for each duck?

If any duck is zero... you may want to talk to the student...

```{r}
duck_avail_heuristics
```

```{r}
# we remove 0 to avoid other problems
indices<-which(duck_avail_heuristics>0)
duck_avail_heuristics_confirmed <- duck_avail_heuristics[indices]
```

transform into a table form

```{r}
duck_avail_heuristics_table = tibble(
  duck = names(duck_avail_heuristics_confirmed),
  avails = duck_avail_heuristics_confirmed
)

duck_avail_heuristics_table
```


another heuristics might be how unique an available slot is?

```{r}
slot_uniqueness_heuristics <- 
 duck_avails %>% 
  select(-day, -bigduck) %>% 
  reframe(timeslot = timeslot, slot_uniqueness = rowSums(across(2:length(.)))) %>% 
  filter(slot_uniqueness > 0) %>% 
  arrange(slot_uniqueness)

slot_uniqueness_heuristics
```

# Make the 1-on-1 meetings

The idea is very simple. Get an available slot, check who is available, schedule it. 
Remove scheduled timeslots and students, and try next one.

```{r}

get_a_schedule = function(max_iter = nrow(duck_ordered_table) * 10, 
                          duck_ordered_table, 
                          slot_order_table,
                          use_duck_order = TRUE){
  meetings <- NULL
  scheduled_ducks <- c()
  iter <- 1
  # if we arranged a meeting for each duck?
  
  while(iter < max_iter && length(scheduled_ducks)!=nrow(duck_ordered_table)){
    
    scheduled_slots <- meetings$timeslot
    
    if(is.null(scheduled_slots)) scheduled_slots = c()
    
    # the first available slot
    current_slot = slot_order_table %>% 
      filter(!(timeslot %in% scheduled_slots))
    # print(current_slot)
    # print(current_slot)
    
    # pull out who is available
    available_ducks = duck_avails %>% 
                      # remove scheduled ducks
                      select(-all_of(scheduled_ducks)) %>% 
                      # who is available
                      filter(timeslot == current_slot$timeslot[1]) %>% 
                      select_if(~ any(. == TRUE)) 
    
   # print(available_ducks)
    
    if(use_duck_order==TRUE){
    # try to schedule more difficult ducks
      next_duck <- duck_ordered_table %>% 
      filter(!(duck %in% scheduled_ducks)) %>% 
      filter(duck %in% colnames(available_ducks)) %>% 
      slice(1) %>% pull(duck)
    }else{
      next_duck <- colnames(available_ducks)[1]
    }
   
    # if this slot is available for next duck
    if(!identical(next_duck, character(0))){
      
      scheduled_ducks <- c(scheduled_ducks, next_duck)
      
      meetings <- meetings %>% 
      rbind(tibble(duck = c(next_duck,next_duck),  
                   timeslot = c(current_slot$timeslot[1], current_slot$timeslot[1] + 1)))
     # print(meetings)
    }
    iter = iter + 1
  }
  if(length(scheduled_ducks)==nrow(duck_ordered_table))
     return(meetings)
  else{
    warning('return a partial schedule.  perhaps try to shuffle around ducks or timeslots')
    return(meetings)
  }
}
```



Deterministic schedule based on heuristics

```{r}
a_schedule = get_a_schedule(duck_ordered_table = duck_avail_heuristics_table,
                            slot_order_table = slot_uniqueness_heuristics)

a_schedule %>% 
  left_join(signup %>% select(timeslot, day, time), by = 'timeslot') %>% 
  arrange(timeslot, duck)
```

So.. if this scheduling isn't too hard. We can just randomize the order and generate another schedule

```{r}
b_schedule = get_a_schedule(duck_ordered_table = duck_avail_heuristics_table[sample(nrow(duck_avail_heuristics_table)),],
                            slot_order_table = slot_uniqueness_heuristics)

b_schedule %>% 
  left_join(signup %>% select(timeslot, day, time), by = 'timeslot') %>% 
  arrange(timeslot, duck)
```

# Evaluate a schedule

We check how many 'Maybe's we have to do. 

```{r}
eval_schedule = function(signup, schedule, message = TRUE){
  
  signup_longer <- signup %>% 
                   pivot_longer(cols = 5:length(.), names_to = 'duck', values_to = 'goodness')
  
  result_goodness <- 
    schedule %>% 
    left_join(signup_longer, by = c('duck', 'timeslot')) %>% 
    arrange(timeslot, duck) 
  
  advisor_col = colnames(result_goodness)[5]
  
  value = mean(result_goodness[[advisor_col]]) + mean(result_goodness$goodness)
  
  if(message){
      cat('1 = Maybe, the advisor\'s maybe:')
      if(nrow(result_goodness  %>% filter(!!sym(advisor_col) != 2)) > 0){
         print(ascii(result_goodness  %>% filter(!!sym(advisor_col) != 2)), type = "rest")
      }else{
         cat('none\n')
      }
      
      cat('1 = Maybe, students\' maybe:')
      if(nrow(result_goodness  %>% filter(goodness != 2)) > 0){
        print(ascii(result_goodness  %>% filter(goodness != 2)), type = "rest")
      }else{
        cat('none\n')
      }
      
      cat(paste0('overall utility is ', value, ' (higher is better) '))
  }
  
  return(value)
}
  
```

```{r}
eval_schedule(signup, a_schedule)
```


```{r}
eval_schedule(signup, b_schedule)
```


# Generate a brunch and get the best

```{r}
seed = seq(10, 100, length.out = 10)
current = 0
schedule = NULL

for(s in seed){
  set.seed(s)
  i_schedule = get_a_schedule(duck_ordered_table = duck_avail_heuristics_table[sample(nrow(duck_avail_heuristics_table)),],
                              slot_order_table = slot_uniqueness_heuristics)
  if(nrow(i_schedule) == 2 * nrow(duck_avail_heuristics_table)){
      #print(i_schedule)
      value = eval_schedule(signup, i_schedule, message = FALSE)
      if(value >= current){
         current = value
         schedule = i_schedule
      }
  }
}
```


```{r}
eval_schedule(signup, schedule)

schedule %>% 
    left_join(signup %>% select(timeslot, day, time), by = 'timeslot') %>% 
    arrange(timeslot, duck)
```

